Rasterizing prodution per each region

### 2022-03-02 analysis/validation_testing.Rmd
cleaned code and updated readme

This script tests our validation set of points against our actual known points.

@@

## establish CRSs

## food 81km2 at equator 
eez_81 <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/data_type_A_eez/eez_validation_mask.tif"))
plot(raster::area(eez_81))
plot(eez_81)
cellStats(raster::area(eez_81), "mean")

test <- data.frame(raster::values(raster::area(eez_81)))

## 36km2 equal area
eez_36_cea <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/data_type_A_eez/eez_validation_mask_cea_36km.tif"))
plot(eez_36_cea)

new_crs <- "+proj=cea +lat_ts=45 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
new_raster <- raster(nrows=3000, ncols=4736, xmn=-14207430, xmx=14208570, ymn=-9000182, ymx=8999818, crs = new_crs)

## 25km2 equal area
eez_25_cea <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/data_type_A_eez/eez_validation_mask_cea_25km.tif"))

cea_25_crs <- crs(eez_25_cea)
plot(eez_25_cea)

## 1/4 degree cells 
eez_quart_cell <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/data_type_A_eez/eez_validation_mask_quarter_degree.tif"))

quarter_crs <- crs(eez_quart_cell)

@@

Read in validation dataset

```{r}
validation_all <- readRDS(file.path("data/all_validation_farms_resubmission.rds"))  %>%
  mutate(species_group = case_when(type == "salmon" ~ "Salmonidae fish",
                          type == "bivalve" ~ "Unfed or algae fed bivalve molluscs",
                          type == "shrimp" ~ "Shrimps and prawns", 
                          type == "tuna" ~ "Bluefin tuna", 
                          type == "marine_fish_general" ~ "General marine fish"))

table_3 <- read_csv("analysis/tables/SI_table_3.csv")

iso <-   data.frame(iso3c = countrycode(table_3$country, origin = 'country.name', destination = 'iso3c'))

table_3_iso3c <- table_3 %>%
  cbind(iso)

data_type_val <- validation_all %>%
  left_join(table_3_iso3c, by = c("iso3c", "species_group"))

data_type_A <- data_type_val %>%
 filter(data_type_2 %in% c("A"))

table_3_A <- table_3_iso3c %>%
  filter(data_type_2 %in% c("A"))

Split data

```{r}
validation_farms <- data_type_A %>%
 filter(source == "modeled validation") %>%
  mutate(count_farms = 1)

val_farms_cea_36 <- validation_farms %>%
  st_transform(crs = new_crs)

val_farms_cea_25 <- validation_farms %>%
  st_transform(crs = cea_25_crs)

val_farms_quarter <- validation_farms %>%
  st_transform(crs = quarter_crs)

real_farms <- data_type_A %>%
  filter(source == "real") %>%
  mutate(count_farms = 1)

real_farms_cea_36 <- real_farms %>%
  st_transform(crs = new_crs)

real_farms_cea_25 <- real_farms %>%
  st_transform(crs = cea_25_crs)

real_farms_quart <- real_farms %>%
  st_transform(crs = quarter_crs)
```

Rasterize data to count number of farms per raster cell with 36km2 equal area

val_rast <- rasterize(val_farms_cea_36, new_raster, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum") # 8552

plot(val_rast, col = "black")
#s <- raster::select(val_rast)
#plot(s)

real_rast <- rasterize(real_farms_cea_36, new_raster, field = "count_farms", fun = "sum")
plot(real_rast)

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8552    8552

Use our EEZ mask to create a confusion matrix for CEA 36km2 cell sizes

 - AUC of ~0.60
```{r}

plot(eez_36_cea, col = "black")

world_coast <- as.data.frame(eez_36_cea) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask_cea_36km))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8535

sum(eez_cells_df$real, na.rm = TRUE) # 8473

## still missing some farms.. but nothing egregious. This is probably ok. The results wouldn't change much even if we had the extra ~80 farms

test <- rast_stack*eez_36_cea

cellStats(test,"sum") 
# layer.1 layer.2 
#    8535    8473

## we are losing some farms...

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

# Confusion Matrix and Statistics
#           Reference
# Prediction       0       1
#          0 1171909    2552
#          1    2721     682
#                                           
#                Accuracy : 0.9955          
#                  95% CI : (0.9954, 0.9956)
#     No Information Rate : 0.9973          
#     P-Value [Acc > NIR] : 1.00000         
#                                           
#                   Kappa : 0.2033          
#                                           
#  Mcnemar's Test P-Value : 0.02069         
#                                           
#             Sensitivity : 0.9977          
#             Specificity : 0.2109          
#          Pos Pred Value : 0.9978          
#          Neg Pred Value : 0.2004          
#              Prevalence : 0.9973          
#          Detection Rate : 0.9949          
#    Detection Prevalence : 0.9971          
#       Balanced Accuracy : 0.6043          
#                                           
#        'Positive' Class : 0

# library(pROC)
# roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE) # 0.6019
# 
# 
# library(pROC)
# library(randomForest)
# 
# 
# roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE, percent = TRUE, xlab = "False Positive Percentage", ylab = "True Positive Percentage", col = "#377eb8", lwd = 4, print.auc = TRUE) # 0.6068
# 
# roc.info <- roc(eez_cells_df$real_count, eez_cells_df$pred_count, legacy.axes = TRUE)
# 
# roc.df <- data.frame(tpp = roc.info$sensitivities*100,
#                      fpp = (1-roc.info$specificities)*100,
#                      thresholds = roc.info$thresholds)
# 
# 
# #### ROC curve try #2
# # https://www.datatechnotes.com/2019/03/how-to-create-roc-curve-in-r.html
# 
# library(ROCR)
#   
#  
# pred = eez_cells_df$pred_count
# 
# pred = prediction(pred, eez_cells_df$real_count)
# 
# perf = performance(pred, "acc")
# plot(perf)
# 
# perf_cost <- performance(pred, "cost")
# plot(perf_cost)
# 
# roc = performance(pred, "tpr", "fpr")
# 
# plot(roc, colorize = T, lwd=2)
# abline(a=0, b=1)
# 
# auc = performance(pred, measure = "auc")
# print(auc@y.values)
# # 0.6091016

Use our EEZ mask to create a confusion matrix for CEA 25km2 cell sizes

 - AUC of ~0.59 - worse than 36km2... makes sense! 
```{r}

val_rast <- rasterize(val_farms_cea_25, eez_25_cea, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum")

plot(val_rast, col = "black")
#s <- raster::select(val_rast)
#plot(s)

real_rast <- rasterize(real_farms_cea_25, eez_25_cea, field = "count_farms", fun = "sum")
plot(real_rast)

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8552    8552

plot(eez_25_cea, col = "black")

world_coast <- as.data.frame(eez_25_cea) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask_cea_25km))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8536

sum(eez_cells_df$real, na.rm = TRUE) # 8482

## still missing some farms.. but nothing egregious. This is probably ok. The results wouldn't change much even if we had the extra ~80 farms. This is likely due to reprojection weirdness

test <- rast_stack*eez_25_cea

cellStats(test,"sum") 
# layer.1 layer.2 
#    8536    8482

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction       0       1
#          0 1689310    2929
#          1    3100     639
#                                           
#                Accuracy : 0.9964          
#                  95% CI : (0.9964, 0.9965)
#     No Information Rate : 0.9979          
#     P-Value [Acc > NIR] : 1.00000         
#                                           
#                   Kappa : 0.1731          
#                                           
#  Mcnemar's Test P-Value : 0.02857         
#                                           
#             Sensitivity : 0.9982          
#             Specificity : 0.1791          
#          Pos Pred Value : 0.9983          
#          Neg Pred Value : 0.1709          
#              Prevalence : 0.9979          
#          Detection Rate : 0.9961          
#    Detection Prevalence : 0.9978          
#       Balanced Accuracy : 0.5886          
#                                           
#        'Positive' Class : 0

Use our EEZ mask to create a confusion matrix for 0.25 X 0.25 degree cell sizes (about 769.3152km2 at equator). Average cell size is 490.3108 km2

 - AUC of ~72.05 ; makes sense that is it better!

```{r}
test <- raster::area(eez_quart_cell)

cellStats(test, "mean") # 490.3108

mean(test@data@values) # 490.3108

val_rast <- rasterize(val_farms_quarter, eez_quart_cell, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum")

plot(val_rast, col = "black")

real_rast <- rasterize(real_farms_quart, eez_quart_cell, field = "count_farms", fun = "sum")
plot(real_rast)

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8552    8552

plot(eez_quart_cell, col = "black")

world_coast <- as.data.frame(eez_quart_cell) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask_quarter_degree))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8124

sum(eez_cells_df$real, na.rm = TRUE) # 8173

## still missing some farms..

test <- rast_stack*eez_quart_cell

cellStats(test,"sum") 
# layer.1 layer.2 
#    8124    8173

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction     0     1
#          0 84570   639
#          1   902   514
#                                           
#                Accuracy : 0.9822          
#                  95% CI : (0.9813, 0.9831)
#     No Information Rate : 0.9867          
#     P-Value [Acc > NIR] : 1               
#                                           
#                   Kappa : 0.3912          
#                                           
#  Mcnemar's Test P-Value : 2.486e-11       
#                                           
#             Sensitivity : 0.9894          
#             Specificity : 0.4458          
#          Pos Pred Value : 0.9925          
#          Neg Pred Value : 0.3630          
#              Prevalence : 0.9867          
#          Detection Rate : 0.9763          
#    Detection Prevalence : 0.9837          
#       Balanced Accuracy : 0.7176          
#                                           
#        'Positive' Class : 0

@@

```{r}
# gentry <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/suitability_rasts/gentry/karakoenig"))
```

>>>>

## Summary

This script tests conducts an accuracy assessment of our model vs the actual, known mariculture locations.

We compare out suitability layer with cells that contain data type A farms. This allows us to see how well our suitability layer overlaps with known, comprehensive farms. In bullet 1, we compare one instance of randomly placing farms into our suitabiltiy layer with data type A farms. I believe if we had ran the model thousands of times, randomly placing data type A farms, it would result in the same conclusion (or something very close) as just comparing our entire suitability layer with data type A farms.

We do these analyses at 4 different resolutions. The native 0.083 degree resolution that we work with all of the global food rasters at. 25km2 and 36km2 equal area, and 0.25 degree resolutions. This allows us to see how well our model works at different resolutions.

@@

@@

## Method: Compare suitability layer to data type A real farms.

@@

### 2022-03-02 marine/compile_data_for_placement.Rmd
cleaned code and updated readme

>>>>

This script compiles all of the location data and the modeling information data into two datasets:

 - all_marine_farms.csv : Lat/Lon locations of all known marine aquaculture farms globally.
 - all_points_to_add.csv : Number of farms per each species group to place in suitable areas, within each country.

### 2022-03-02 marine/last_step_rasterize/shrimp_rasterize_production.Rmd
cleaned code and updated readme

shrimp_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)
plot(log(shrimp_raster + 1))

@@

>>>>

shrimp_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)
plot(log(shrimp_raster + 1))
cellStats(shrimp_raster, "sum")

@@

writeRaster(shrimp_raster, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/production/rasters/shrimp_production_rast.tif")
```
